```{r}
library(rTensor)
library(quantmod)
```

```{r}
tech_tickers <- c("AAPL", "MSFT", "GOOG", "NVDA", "META", "AMZN", "TSLA", "IBM", "ORCL", "INTC", "NFLX", "PYPL", "ADBE", "CSCO", "QCOM", "TXN", "AMD")
energy_tickers <- c("XOM", "CVX", "NEE", "SLB", "COP", "EOG", "OXY", "KMI", "VLO", "PSX", "WMB", "LNG", "MPC", "HES", "APA", "TRP", "VST", "FANG")
financial_tickers <- c("JPM", "BAC", "GS", "MS", "WFC", "C", "AXP", "BLK", "TROW", "SCHW", "PRU", "CME", "MET", "COF", "AIG", "SPGI", "TFC", "TRV")

combined_tickers <- c(tech_tickers, energy_tickers, financial_tickers)

# downloading stock data using the quantmod library
getSymbols(combined_tickers, src = "yahoo", from = "2023-01-01", to = "2023-12-31")
```

We now form our tensor by stacking the dataframes.

```{r}
stockdata <- array(NA, dim = c(length(combined_tickers), 6, 250))

for(i in 1:length(combined_tickers)){
  stockdata[i,,] <- t(get(combined_tickers[i]))
}
stockdata_tensor <- as.tensor(stockdata)
```

Compute the CP decomposition followed by k-means on on the first factor matrix

```{r}
set.seed(2025)

for(i in 1:dim(stockdata_tensor)[1]){
  for(j in 1:dim(stockdata_tensor)[2]){
    stockdata_tensor[i,j, ] <- scale(stockdata_tensor[i,j, ]@data)
  }
}

cp_stocks <- cp(stockdata_tensor, num_components = 14, max_iter = 1000) # 11 and 14 components work best

# extract factor matrices from the CP output
U1_cp_stocks <- cp_stocks$U[[1]]
U2_cp_stocks <- cp_stocks$U[[2]]
U3_cp_stocks <- cp_stocks$U[[3]]

colouring <- c(rep("red", length(tech_tickers)), rep("blue", length(energy_tickers)), rep("green", length(financial_tickers)))

pairs(U1_cp_stocks, col=colouring, pch=19) # colour based on industry

kmeans_res <- kmeans(U1_cp_stocks, centers = 3)

table(Predicted = kmeans_res$cluster, Actual = colouring)

plot(U1_cp_stocks[,1], U1_cp_stocks[,6], col=colouring, pch=19, xlab = "First column of U1", ylab = "Sixth column of U1")
legend("topleft", legend = c("Technology", "Financial", "Energy"), col = c("red", "green", "blue"), pch=19, cex = 1.5)
```
Now, we unfold and conduct PCA, followed by a k-means on the first five PCs.

```{r}
set.seed(1)

pca_compare <- prcomp(k_unfold(stockdata_tensor, 1)@data, center = TRUE, scale. = TRUE)

pc_result <- pca_compare$x[,1:5]

pairs(pc_result, col=colouring, pch=19)

kmeans_pca <- kmeans(pc_result, centers = 3)

# confusion matrix
table(Predicted = kmeans_pca$cluster, Actual = colouring)
```
Tech stocks are on average trending upwards.

```{r}
par(mfrow=c(3,3))
plot(AAPL$AAPL.Open)
plot(MSFT$MSFT.Open)
plot(TSLA$TSLA.Open)
plot(META$META.Open)
plot(AMZN$AMZN.Open)
plot(ORCL$ORCL.Open)
plot(NFLX$NFLX.Open)
plot(ADBE$ADBE.Open)
plot(CSCO$CSCO.Open)
```

Energy stocks show upward trend around July.

```{r}
par(mfrow=c(3,3))
plot(XOM$XOM.Open)
plot(FANG$FANG.Open)
plot(APA$APA.Open)
plot(SLB$SLB.Open)
plot(COP$COP.Open)
plot(EOG$EOG.Open)
plot(OXY$OXY.Open)
plot(KMI$KMI.Open)
plot(VLO$VLO.Open)
```

Financial stocks are somewhat steady with a rally towards the end of the year.

```{r}
par(mfrow=c(3,3))
plot(JPM$JPM.Open)
plot(BAC$BAC.Open)
plot(GS$GS.Open)
plot(MS$MS.Open)
plot(WFC$WFC.Open)
plot(C$C.Open)
plot(AXP$AXP.Open)
plot(TROW$TROW.Open)
plot(MET$MET.Open)
```

Plotting the first and the third column of the second factor matrix.

```{r}
par(mfrow=c(1,2))
plot(U2_cp_stocks[,1], ylab="First Column of U2", pch=19)
plot(U2_cp_stocks[,3], ylab="Third Column of U2", pch=19)
```

